---
title: "Activity Summary"
output: html_document
---


### Read in the Data


```{r}
act <- read.csv("activity.csv", na.strings = "NA", stringsAsFactors = FALSE, colClasses = c("numeric","Date","numeric"))
```

## What is mean total number of steps taken per day?


1) Prepare data for analysis

```{r}
act1 <- na.omit(act)
act1_byday <- aggregate(act1$steps, by=list(day=act1$date), FUN=sum)
```

2) Plot the histogram

```{r echo = TRUE}
hist(act1_byday$x, breaks=59, main="Total Steps Taken Per Day", xlab = "Steps")
```

3) Calculate the mean

```{r echo = TRUE}
mean(act1_byday$x)
```

4) Calucluate the median

```{r echo = TRUE}
median(act1_byday$x)
```



## What is the average daily activity pattern?


1) Prepare data for analysis and plot

```{r echo=TRUE}
act2 <- aggregate(steps ~ interval, data = act, mean, na.rm = TRUE)
plot(act2$interval, act2$steps, type = "l", lwd = 2, col = "green", main = "Average Number of Steps Taken", xlab = "5-Minute Interval", ylab = "Average Steps")
```


2) Calculate 5-minute interval with maximum number of steps

```{r echo=TRUE}
act2$interval[which.max(act2$steps)]
```


## Imputing missing values


1) Calculate number of missing NAs

```{r echo=TRUE}
table(complete.cases(act))
```

As we are calculating NAs across any row in any column, we use the "FALSE" count as the total, therefore the answer is 2,304.


2) Create a way to fill in the missing values. Do this by looping through the full "activity" dataset and filling in the missing values with the average values of each internal (from question above)

```{r echo=TRUE}
act3 <- act 
for (i in act2$interval) {
        act3[act3$interval == i & is.na(act3$steps), ]$steps <- act2$steps[act2$interval == i]
}
table(complete.cases(act3))
```

Using the table command again, we see that there are not FALSE values, hence no "incomplete"" cases

3) Plot and calculate mean and median as similar to question #1 above


```{r echo=TRUE}
act3_byday <- aggregate(act3$steps, by=list(day=act3$date), FUN=sum)

hist(act3_byday$x, breaks=59, main="Total Steps Taken Per Day", xlab = "Steps")

mean(act3_byday$x)
median(act3_byday$x)
```

There really isn't much impact to the resulting data from imputing the values. This isn't really surprising considering that the imputed values were averages of existing data. Overall, there is very little impact from the imputed data.


## Are there differences in activity patterns between weekdays and weekends?


1) Create dataset with "weekdays" factor

```{r echo=TRUE}
act4 <- act3
act4$weekday <- weekdays(act4$date, abbreviate = FALSE)

act4$weekpart <- ""

act4[act4$weekday == "Saturday" | act4$weekday == "Sunday", ]$weekpart <- "Weekend"
act4[!(act4$weekday == "Saturday" | act4$weekday == "Sunday"), ]$weekpart <- "Weekday"
```

2) Plot the average 5-minute interval for each "weekday" or "weekend"


```{r echo=TRUE}
library(lattice)

act4_byday <- aggregate(steps ~ interval + weekpart, data = act4, mean)

xyplot(steps ~ interval | weekpart, data = act4_byday, layout = c(1, 2), type="l")

```

